---
title: "Course 8 - Practical Machine Learning - Prediction Assignment"
author: "Pedro M Batista"
date: "12/10/2018"
output: html_document
---

Summary
1. Introduction
2. Packages required
3. Getting Data
4. Dissecting the Data
5. Prediction Modelling
5.i. Decision Tree
5.ii. Random Forest Model
6. Predicting with Random Forest Model
7. Choosing the Model in Predicting to test the data

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## 1. Introduction
We aim at predicting how our subjects exercise. In this report we describe how we built our model and how we used cross validation.

## 2. Packages required

```{r, echo=FALSE}

library(rpart.plot)
library(rpart)
library(RColorBrewer)
library(randomForest)
library(gbm)
library(plyr)
library(caret)
library(knitr)
library(caret)
library(rattle)
library(corrplot)

```

## 3. Getting Data

```{r, echo=FALSE}
download.file(url = "https://d396qusza40orc.cloudfront.net/predmachlearn/pml-training.csv", destfile = "./pml-training.csv")
dtraining <- read.csv("./pml-training.csv", na.strings=c("NA","#DIV/0!",""))
download.file(url = "https://d396qusza40orc.cloudfront.net/predmachlearn/pml-testing.csv", destfile = "./pml-testing.csv")
dtesting <- read.csv("./pml-testing.csv", na.strings=c("NA","#DIV/0!",""))

```

## 4. Dissecting the Data

We partition the data into training (60%) and testing (40%).

```{r, echo=FALSE}
valid <- names(dtesting[,colSums(is.na(dtesting)) == 0])[8:59]
dtraining <- dtraining[,c(valid,"classe")]
dtesting <- dtesting[,c(valid,"problem_id")]
dim(dtraining); dim(dtesting);
```

```{r}
set.seed(12345)

Train <- createDataPartition(dtraining$classe, p=0.55, list=FALSE)
training <- dtraining[Train,]
testing <- dtraining[-Train,]

dim(training); dim(testing);
```

## 5. Prediction Modelling

Now we build a (i) Decison Tree and (ii) Random Forest Model.
  
## 5.i. Decision Tree

```{r}
set.seed(301)
modFitDecTree <- rpart(classe ~ ., data=training, method="class")
fancyRpartPlot(modFitDecTree)
```
```{r, echo=FALSE}
predictTree <- predict(modFitDecTree, newdata=testing, type="class")
confTree <- confusionMatrix(predictTree, testing$classe)
confTree

plot(confTree$table, col = confTree$byClass, 
     main = paste("Decision Tree - Accuracy =",
                  round(confTree$overall['Accuracy'], 4)))
```

## 5.ii. Random Forest Model

The error is estimated with 40% testing sample.

```{r}
set.seed(12345)

modFitRF <- randomForest(classe ~ ., data = training, method = "rf", importance = T, trControl = trainControl(method = "cv", classProbs=TRUE,savePredictions=TRUE,allowParallel=TRUE, number = 10))

plot(modFitRF)
```

## 6. Predicting with Random Forest Model
```{r}
prediction <- predict(modFitRF, testing, type = "class")
confusionMatrix(prediction, testing$classe)
```

Random forest model obtains 99.3% Accuracy.

## 7. Choosing the Model in Predicting to test the data

Now we apply the random forest model with an accuracy of 99.3% over Decision tree:

```{r}
predictTEST <- predict(modFitRF, newdata=testing)
predictTEST
```

